\hspace{1cm} Do syntactic and semantic similarity lead to interference effects? \newline Evidence from self-paced reading and event-related potentials using German

### 2024-01-04 paper/main.tex
minor edits in conclusion

\section{Conclusion}
This project investigated the use of  syntactic and semantic cues during subject-verb dependency resolution with self-paced reading and event-related potentials.  Both our experiments consistently showed semantic interference, i.e., more processing difficulty during long-distance subject-verb dependency resolution in the presence of a semantically matching distractor. This manifested in longer self-paced reading times starting at the distractor and continuing into the post-critical region, and in a more negative N400 amplitude for the high vs.\ low semantic interference conditions elicited by the critical verb. These different time courses in the two experiments are likely to be due to a mix of encoding and retrieval interference.

The observed differences in the time course in the SPR and ERP studies are likely to be due to the methods used in the two sets of experiments (self-paced vs.\ auto-paced reading). Overall, we found evidence against the use of a syntactic cue such as \{$\pm$ grammatical subject\}, as previously assumed in the literature; computational modeling shows that--at least given the present data--the parser uses the syntactic cue  \{$\pm$ subject-in-same-clause\}  to identify the correct target for retrieval.  A broader implication is that the parser may be able to use hierarchical structure in the input to target the correct syntactic dependent from memory, leading to no syntactic interference from a distractor that is in subject position within an embedded clause.

>>>>

\section{Conclusion}
This project investigated the use of  syntactic and semantic cues during subject-verb dependency resolution with self-paced reading and event-related potentials.  Both our experiments consistently showed semantic interference, i.e., more processing difficulty during long-distance subject-verb dependency resolution in the presence of a semantically matching distractor. This manifested in longer self-paced reading times starting at the distractor and continuing into the post-critical region, and in a more negative N400 amplitude for the high vs.\ low semantic interference conditions at the critical verb. The observed differences in the time course of the effect are likely to be due to the methods used in the two sets of experiments (self-paced vs.\ auto-paced reading), which might have lead to a different mixture of encoding and retrieval interference. Overall, we found evidence against the use of a syntactic cue such as \{$\pm$ grammatical subject\}, as previously assumed in the literature; computational modeling shows that -- at least given the present data -- the parser uses the syntactic cue  \{$\pm$ subject-in-same-clause\}  to identify the correct target for retrieval.  A broader implication is that the parser may be able to use hierarchical structure in the input to target the correct syntactic dependent from memory, leading to no syntactic interference from a distractor that occurred within an embedded clause.
